Iroko: Predictive Datacenter Congestion Control

### 2017-12-12 Documents/Report/conclusion.tex
Edits

Iroko represents a first attempt at using machine learning
to predict data center traffic. The endeavour relies crucially
on data center traffic having enough structure to in fact allow
some level of prediction. Yet even for largely random traffic
Iroko showed decent results, often besting baseline solutions.
Iroko is only the opening salvo, but the attempt has merit. 
More work must be done before claims of superiority to all
incumbent solutions can be made, but there are reasons to be
optimistic.

Machine learning is perhaps the king of "black-box" solutions, 
and it's efficacy is undeniable if sometimes misleading. While
the standoff between so called "white-box" and "black-box" is 
likely to continue for the foreseeable future, Iroko shows
that "black-box" solutions do have a role to play. It may turn
out that neither comes to rule, but rather future SDN's will 
employ both techniques to build ever more efficient solutions.
It may well turn out that the future is grey.

>>>>

Iroko represents a first attempt at using machine learning
to predict datacenter traffic. The endeavour relies crucially
on datacenter traffic having enough structure to in fact allow
some level of prediction. Yet even for largely random traffic,
Iroko showed decent results, often besting baseline solutions.
For more predictable traffic, Iroko can compete with "best-in-class"
solutions, and at times shows the strongest performance of all the
tested solutions.
Iroko is only the opening salvo, but the attempt has merit. Iroko
struggles with reducing drop rate, and some random traffic
patterns. With more fine-tuned
prediction, and better rate-limiting algorithms, Iroko can likely
overcome these challenges. 
More work must be done before claims of superiority to all
incumbent solutions can be made, but there are reasons to be
optimistic.

Machine learning is perhaps the king of "black-box" solutions, 
and its efficacy is undeniable, if sometimes misleading. While
the standoff between so called "white-box" and "black-box" solutions is 
likely to continue for the foreseeable future, Iroko shows
that "black-box" solutions do have a role to play. It might be the case
that neither comes to rule, but rather future SDN's will 
employ both techniques to build ever more efficient solutions.
It may well turn out that the future is grey.

### 2017-12-12 Documents/Report/design.tex
Edits

\subsection{Overview and Goals}
In our initial system, "Iroko", a centralized controller regulates all 
node traffic by rate-limiting end-hosts. We have opted for a centralized 
manager in favor of a distributed protocol to leverage the advantages of a 
global network view.~\footnote{For now, we do not expect our design to scale up 
to thousands of switches. Iroko is intended for small to mid-tier size data 
centers, which may benefit from a simplistic, centralized scheduling model.}
By polling each switch individually for port and utilization statistics we are 
able to infer a global traffic matrix, which we can amalgamate with static 
routing and topology information.

@@

The controller will dynamically adjust bandwidth allocations to ensure that hosts 
requiring more bandwidth are allocated it, while host that are not currently 
utilizing all their allocated bandwidth have their allocated bandwidth reduced. 
To avoid starving hosts, some small amount of bandwidth must always be 
allocated to a given host even if it not utilizing network resources. Thus, the 
network, by design, will never reach 100\% utilization.

Conversely however one should rarely see dropped packets due to network 
congestion. Furthermore, the controller must be able to react quickly to 
changing bandwidth needs for hosts. Ideally, using a predictive method.

@@

The environment in this case is defined as the statistics collected from each 
switch and the current end-host to rate limit. This decision is based on the fact that these data are easy to collect and provides a current snapshot of the performance of the data center. Using this representation, we take full advantage of all the information provided in our data center.

We wish to optimize over the packet loss, and define the reward to be good (+1) if there is no loss and bad (-1) if there is any loss.  One limitation of the current reward signal is it's inability to distinguish cases when loss is 0 because no bandwidth was used in the epoch. This means, a more intricate reward functions can be devised for  a future work.

Given the data center assumption, the number of end-hosts is known in the 
topology. We define our arbiter's actions as a single continuous value which will either increase 
decrease, or maintain the allocated bandwidth to different end hosts. A continuous value allows our agent to make direct rate-limiting decisions, and reinforcement learning algorithms have been devised to learn continuous value actions such as this ~\cite{DDPG}.

Thus we frame our problem as one which can now be solved using techniques in reinforcement learning. Given full representation of our environment and 
actions, we will store the state-action value, also referred to as the Q(s,a) function,  in an artificial neural network. The Q(s,a) function 
stores the reward signal information that is used to make future action choices. We chose an artificial neural network to approximate this function, because of they are theoretically proven function approximators ~\cite{Cybenko1989} which have made intractable environmental representations manageable~\cite{Sutton:1998:IRL:551283}.

@@

When a node opens a flow, it will look up the current bandwidth restriction for 
the destination IP in its token database and calculate the maximum congestion 
window possible for the particular TCP connection. In our simple design, flows 
are always assigned a fair fraction.

@@

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{Topology3}
\caption{Simple sketch of the Iroko architecture. A central scheduler maintains 
a consistent view of the current network activity and computes the optimal 
end-host bandwidth distribution. Bandwidth is enforced at end-hosts on a per-IP 
basis.}
\label{fig:Topology3}
\end{figure}

>>>>

\subsection{Overview and Goals}
In our initial system, "Iroko", a centralized controller regulates all 
node traffic by rate-limiting end-hosts. We have opted for a centralized 
manager in favor of a distributed protocol to leverage the advantages of a 
global network view.~\footnote{For now, we do not expect our design to scale up 
to thousands of switches. Iroko is intended for small to mid-tier size datacenters,
which may benefit from a simplistic, centralized scheduling model.}
By polling each switch individually for port and utilization statistics we are 
able to infer a global traffic matrix, which we can amalgamate with static 
routing and topology information.

@@

The controller will dynamically adjust bandwidth allocations to ensure that hosts 
requiring more bandwidth are allocated it, while host that are not currently 
utilizing all their allocated bandwidth have their allocated bandwidth reduced. 
To avoid starving hosts, some small amount of bandwidth must always be 
allocated to a given host even if it not utilizing network resources. Thus, the 
network, by design, will never reach 100\% utilization. Conversely, one should rarely see
dropped packets due to network 
congestion. Furthermore, the controller must be able to react quickly to 
changing bandwidth needs for hosts, ideally, using a predictive method.

@@

The environment in this case is defined as the statistics collected from each 
switch and the current end-host to rate limit. This decision is based on the fact that these data are easy to collect and provides a current snapshot of the performance of the datacenter. Using this representation, we take full advantage of all the information provided in our datacenter.

We wish to optimize over the packet loss, and define the reward to be good (+1) if there is no loss and bad (-1) if there is any loss.  One limitation of the current reward signal is it's inability to distinguish cases when loss is 0 because no bandwidth was used in the epoch. This means, more intricate reward functions can be devised for  a future work.

Given the datacenter assumption, the number of end-hosts is known in the 
topology. We define our arbiter's actions as a single continuous value which will either increase 
decrease, or maintain the allocated bandwidth to different end hosts. A continuous value allows our agent to make direct rate-limiting decisions, and reinforcement learning algorithms have been devised to learn continuous value actions such as this ~\cite{DDPG}.

Thus we frame our problem as one which can now be solved using techniques in reinforcement learning. Given full representation of our environment and 
actions, we will store the state-action value, also referred to as the Q(s,a) function,  in an artificial neural network. The Q(s,a) function 
stores the reward signal information that is used to make future action choices. We chose an artificial neural network to approximate this function. Neural networks are theoretically proven function
approximators ~\cite{Cybenko1989} which have made intractable environmental representations manageable~\cite{Sutton:1998:IRL:551283}.

@@

When a node opens a flow, it looks up the current bandwidth restriction for 
the destination IP in its token database and calculates the maximum congestion 
window possible for the particular TCP connection. In our simple design, flows 
are always assigned a fair fraction.

@@

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{Iroko}
\caption{Simple sketch of the Iroko architecture. A central scheduler maintains 
a consistent view of the current network activity and computes the optimal 
end-host bandwidth distribution. Bandwidth is enforced at end-hosts on a per-IP 
basis.}
\label{fig:Topology3}
\end{figure}

### 2017-12-12 Documents/Report/discussion.tex
Edits

Rather predictably, Iroko struggles on random traffic patterns, however we
do not see this as particularly concerning. Firstly \textit{Iroko} still outperforms
both DCTCP and ECMP implementations. Further, as these traffic patterns
were designed by the authors of the Hedera paper \cite{hedera}, it remains
likely these patterns are somewhat amenable to the Hedera implementation.
Finally random traffic patterns seem unlikely to be prevalent in a
datacenter context. While the general predictability of datacenter
traffic remains an area of open research, our results show that
if there is sufficient predictability, we do get results better than
random, and for truly random data, Iroko does no worse generally than
incumbent solutions.

Our machine learning approach shows promise, but machine learning remains
largely a black-box discipline. This nature makes it difficult for us
to determine precisely which parameters to use, and what drives
the improvements we've seen. Our current machine learning approach
is remarkably light-weight and does not require vast resources
in our test network. A heavier handed approach may yield better
results, however these may come at the cost of scalability.

>>>>

Not unexpectedly, Iroko struggles on a few random traffic patterns, however we
do not see this as particularly concerning. Firstly, Iroko still outperforms
both DCTCP and ECMP implementations on most random traffic patterns, and in general.
Further, as these traffic patterns
were designed by the authors of the Hedera paper \cite{hedera}, it remains
likely these patterns are somewhat amenable to the Hedera implementation.
Finally, as discussed, fully random traffic patterns seem unlikely to be prevalent in a
datacenter context. While the general predictability of datacenter
traffic remains an area of open research, our results show that
if there is sufficient predictability, we do get results better than
random, and for truly random data, Iroko does no worse generally than
incumbent solutions.

Our machine learning approach shows promise, but machine learning remains
largely a "black-box" discipline. This nature makes it difficult for us
to determine precisely which parameters to use, and what drives
the improvements we've seen. Our current machine learning approach
is remarkably light-weight and does not require vast resources
in our test network. A heavier handed approach may yield better
results, however these may come at the cost of scalability. We
also note that further tuning of the current implementation is
likely to lead to stronger results. Due to testing limitations, 
the controller had a very short number of periods to learn 
on. In longer running cases, we fully expect Iroko to 
deliver stronger results as there is more time for machine learning 
algorithms to converge.

For efficient routing, we used EMCP with Iroko, however better
routing regimens exist, in particular Hedera. While we compared
Iroko to Hedera, the two systems are largely orthogonal
and may in fact benefit from one another. Iroko does not directly deal
with routing or so called elephant flows. Thus, Iroko could be
further improved by employing a routing scheme similar to that of
Hedera, while still retaining bandwidth prediction capabilities.

### 2017-12-12 Documents/Report/eval.tex
Edits

This section describes the evaluation of Iroko using our testbed built on top of
Mininet simulator. The main goal is to compare the performance against the
existing centralized and decentralized techniques like Hedera~\cite{hedera} and
DCTCP~\cite{dctcp} respectively.

@@

For evaluation purposes we are targeting a small scale data center of 16 hosts
and 20 switches with FatTree topology. We took measurements for Hedera, DCTCP,
ECMP and with Non-blocking switch on the same test setup. ECMP measurement is
considered as the base line. Non-blocking is a hypothetical single switch fully
connected setup and is considered as the theoretical maximum we are aiming to
achieve.

For traffic simulation, we are using the same traffic matrices used in
Hedera~\cite{hedera}. We took measurements for the variations of the following 3
traffic patterns:

\begin{enumerate} \item Stride$\left(i\right)$: A host with index $x$ sends to
the host with index $(x + i)mod(num hosts)$.  \item Staggered Prob $\left(EdgeP,
PodP\right)$: A host sends to another host in the same edge switch with
probability $EdgeP$, and to its same pod with probability $PodP$, and to the
rest of the network with probability $1-EdgeP - PodP$.  \item Random: A host
sends to any other host in the	network with uniform probability. We include
bijective mappings and ones where hotspots are present.  \end{enumerate}

Since we are primarily concerned about reducing the latency and packet drops
while keeping utilization at maximum, we measured bisection bandwidth and
    average queue length on switches. The experiments last for 60 seconds for
    each benchmark and we consider only the middle 40 seconds.

@@

\begin{figure*} 
	\centering 
	\includegraphics[width=1\linewidth]{rate.png} 
	\caption{}
	\label{fig:rate} 
\end{figure*}

In Figure \ref{fig:rate} we show the aggregate bisection bandwidth. Iroko
consistently provides better utilization compared to ECMP and DCTCP in almost
all experiments. As expected non-blocking is providing the best performance out
of all four techniques. Utilization of non-blocking is lower for traffic
patterns with high incast. Hedera perform extremely well in all experiments
except for some where Iroko outperforms Hedera. Interestingly enough, Hedera
even outperforms non-blocking limit in stride8 experiment. ECMP does reasonably
well when there is high probability of local communication for example smaller
strides, but start to deteriorate as the stride length increases because of hash
collisions.

Another interesting observation is that DCTCP gives abysmal utilization for some
experiments. Further investigation into this revealed that this is because of
DCTCP's aggressive congestion window size reduction strategy on incast. The
proposed solution for incast problem is to use dynamic buffer allocation in
switches which is not implemented in our test setup. General performance of
DCTCP is not satisfactory either. For some experiments it even under performs
ECMP.  According to DCTCP paper it is optimized for some specific data center
traffic patterns like web search, soft real-time applications, recommendation
systems etc which we don't include in our experiments.

@@

Apart from utilization and loss rate following are the other benchmarks that we
are interested. We have not measured these yet due to time concerns.

@@

\item \textbf{99th percentile latency:} Since we are aiming for a low latency
which means that 99th percentile latency in the network across all flows should
be as low as possible.

\item \textbf{Fairness:} Fairness across flows can be tested by introducing a
new host to a completely saturated network and increasing the transmission rate
to see if all the hosts gets fair share of the total bandwidth. Currently we are
assuming all flows should be at equal priority. Differential priority is out of
scope.

>>>>

This section describes the evaluation of Iroko using our testbed built on the
Mininet simulator. The main goal is to compare the performance against the
existing centralized and decentralized techniques like Hedera~\cite{hedera} and
DCTCP~\cite{dctcp}.

@@

For evaluation purposes, we are targeting a small scale data center of 16 hosts
and 20 switches with a FatTree topology. We take measurements for Hedera, DCTCP,
ECMP and a non-blocking switch on the same test setup. We consider the ECMP measurement
to be the base line. The non-blocking result represents a hypothetical single switch with full
backplane bandwidth, and is considered the theoretically achievable maximum.

For traffic simulation, we use the same traffic matrices used in
Hedera~\cite{hedera}. We measured the following 3
traffic patterns:

\begin{enumerate} \item Stride$\left(i\right)$: A host with index $x$ sends to
the host with index $(x + i)mod(num hosts)$.  \item Staggered Prob $\left(EdgeP,
PodP\right)$: A host sends to another host on the same edge switch with
probability $EdgeP$, and to its same pod with probability $PodP$, and to the
rest of the network with probability $1-EdgeP - PodP$.  \item Random: A host
sends to any other host in the	network with uniform probability. We include
bijective mappings and ones where hotspots are present.  \end{enumerate}

Since we are primarily concerned about reducing the latency and packet drops
while keeping utilization at maximum, we measured bisection bandwidth and
    average queue length on the switches. The experiments lasted for 60 seconds for
    each benchmark and we consider only the middle 40 seconds.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{rate.png}
	\caption{Average bisection bandwidth}
	\label{fig:rate}
\end{figure*}

In Figure \ref{fig:rate} we show the aggregate bisection bandwidth. Iroko
consistently provides better utilization compared to ECMP and DCTCP in almost
all experiments. As expected, non-blocking is providing the best performance out
of all four techniques. Utilization of non-blocking is lower for traffic
patterns with high incast. Hedera performs extremely well in all experiments
except for a few like stride1, stag2$(0.5,0.3)$ where Iroko outperforms Hedera.
Interestingly enough, Hedera even outperforms non-blocking in the stride8
experiment. ECMP does reasonably well when there is high probability of local
communication; for example smaller strides, but starts to deteriorate as the
stride length increases because of hash collisions.

Another interesting observation is that DCTCP gives abysmal utilization for some
experiments, for example, randx2 and randx4. Further investigation into this
revealed that this is because of DCTCP's aggressive congestion window size
reduction strategy on incast. The proposed solution for incast problems is to use
dynamic buffer allocation in switches, which is not implemented in our test
setup. General performance of DCTCP is not satisfactory either. For some
experiments it even under-performs ECMP.  According to the DCTCP paper, it is
optimized for some specific data center traffic patterns like web search, soft
real-time applications, recommendation systems etc., which we don't include in our
experiments.

\subsection{Drop rate}

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{qlen.png}
	\caption{CDF of queue lengths}
	\label{fig:qlen}
\end{figure*}

To analyze the drop rate we measured the number of packets built up in the queue
on all switches for the same set of traffic patterns and plotted a cumulative
distribution graph. In Figure \ref{fig:qlen} shows some selected
graphs. Recall that higher queue utilization is considered less optimal. For our experiments, we fixed the maximum
queue size at 50 packets.

Iroko struggles with queue management; consistently keeping $90\%$ of the queues
empty while the remaining $10\%$ are full. It is worth noting however, that Iroko performs
very well for traffic patterns like stride1, stag2$(0.5,0.3)$ etc., where it
shows high utilization. DCTCP maintains the queues at a minimum across
all experiments. This is not surprising considering the low utilization of
DCTCP. ECMP and Hedera behave almost the same. Non-blocking has the highest
queue size out of all techniques. We surmise that the CPU is the bottleneck here;
non-blocking uses only a single switch to connect the entire network with full
backplane bandwidth.

Iroko could definitely benefit from more fine tuning for the prediction algorithm. We further
need to investigate why the CDF of queue lengths exhibits a sudden jump from $0$ to $50$. Initial
analysis shows that a small fixed set of switches are the bottleneck in all
experiments. Our guess is that this happens because of the unfair bandwidth
allocation across traffic flows. Analyzing fairness might provide better insight
into this problem.

@@

Apart from utilization and loss rate, the following benchmarks remain of interest.
Due to time constraints, we were not able to fully investigate these.

@@

\item \textbf{99th percentile latency:} We are aiming for a low latency solution,
which means that the 99th percentile latency in the network across all flows should
be as low as possible.

\item \textbf{Fairness:} Fairness across flows can be tested by introducing a
new host to a completely saturated network and increasing the transmission rate
to see if all the hosts gets fair share of the total bandwidth. Currently we are
assuming that all flows should be given equal priority. Differential priority is out of
scope.

### 2017-12-12 Documents/Report/future_work.tex
Edits

Predicting datacenter traffic online with machine learning shows promise, and we believe that 
future widely deployed datacenter traffic control mechanisms will include
machine learning in their toolkit. However machine learning alone may not
be sufficient to achieve the best possible results. Machine learning
based central traffic controllers can benefit from dynamic flow
scheduling as well as further traffic engineering. Thus expanding
the Iroko framework to include flow rather than interface based 
predictions, or a combination of both, would be an interesting 
avenue to pursue. Combined with a robust routing algorithm, we
believe such a solution show superior performance when
compared to virtual all currently existing solutions.

For large, long-running workloads that require significant 
network resources at the termination such as MAP-REDUCE
workloads, hosts may have very predictable completion times
for these jobs. In such cases, hosts could report these 
estimated completion times to the central controller,
allowing the controller to account for them pre-emptively
rather than trying to infer them which may take a couple
of epochs, and therefore reduce efficiency. Such a
system would be complementary to Iroko and
allow for finer grained control. By taking further
host level input, Iroko could reason about individual 
flows, and allocate not only bandwidth, but also 
specific paths.

>>>>

Predicting datacenter traffic online with machine learning shows promise, and we believe that 
future widely deployed datacenter traffic control mechanisms will include
machine learning in their toolkit. However, machine learning alone may not
be sufficient to achieve the best possible results. Machine learning
based central traffic controllers can benefit from dynamic flow
scheduling as well as further traffic engineering. Thus expanding
the Iroko framework to include flow rather than interface based 
predictions, or a combination of both, would be an interesting 
avenue to pursue. Combined with a robust routing algorithm, we
believe such a solution could show superior performance when
compared to virtually all currently existing solutions.

For large, long-running workloads that require significant 
network resources at the termination such as MAP-REDUCE
workloads, hosts may have very predictable completion times
for these jobs. In such cases, hosts could report these 
estimated completion times to the central controller,
allowing the controller to account for them pre-emptively
rather than trying to infer them, which may take a couple
of epochs, and therefore reduce efficiency. Such a
system would be complementary to Iroko and
allow for finer grained control. By taking further
host level input, Iroko could reason about individual 
flows, and allocate not only bandwidth, but also 
specific paths.

### 2017-12-12 Documents/Report/implementation.tex
Edits

\subsection{Topology Simulation}
We chose to use  Mininet, a python data center emulation API, to prototype Iroko. Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}. This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware. Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot be deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.

In the Mininet environment, we created a fat-tree topology consisting of 20 switches and 16 hosts . The 20 switches are configured into 5 pods as described in ~\cite{fattree}. This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows. In ~\cite{hedera}, Hedera was evaluated  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .

Mininet’s lightweight virtualization also allows the API to scale such as to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}. This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.

@@

In addition, we used the following Linux command:\\
\texttt{tc -s qdisc show dev <interface>}\\
where \texttt{<interface>} is the current host we are polling from. Using the 
obtained text, we then applied regex expressions to collect the packet loss, 
packet overlimits, and interface queues. Each of these data provide further 
information to hint at the current state of the data center. Our goal was to 
eliminate dropped packets, and used this to help Iroko make rate-limiting 
decisions. Overlimit packets provides information on when a link goes over it’s 
allocated bandwidth, and adds a potentially useful feature for our controller 
to learn from. Interface queues provide Iroko with information on the queue 
length of the host, which measures how many packets are buffered to be 
processed.  These additional statistics we hoped would provide our controller 
enough data to make meaningful inferences of the appropriate bandwidth 
allocation to each host.

\subsection{Bandwidth Allocation}
Iroko rate limits traffic at the applications level, and thus uses UDP instead of TCP to communicate with hosts. It reads a traffic matrix shared among all the hosts and sends network traffic accordingly. Iroko also manages the current allocated bandwidth per host by sending a packet which updates their current allocated bandwidth.

@@

We chose to use the deep deterministic policy gradient policy algorithm (DDPG) almost entirely as described in ~\cite{DDPG} . DDPG is a form of actor-critic reinforcement learning algorithm ~\cite{Sutton:1998:IRL:551283} which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.  Our agent learns using replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}. This decision might be unnecessary for a data center where authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and this suggests that we could assume data center data is inherently uncorrelated. We recognize these findings may also give an upper limit to the expected performance of Iroko.

To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}. Instead of the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$ ~\cite{Sutton:1998:IRL:551283}. Our random action is then defined as follows:

@@

uniform(0,1), &\text{P(x = increase) = .333},\\
uniform(-1,0), &\text{P(x = decrease) = .3333},\\
0, & \text{(x = unchanged) = .3333}
\end{cases}
\]

where uniform(a,b) refers to taking a random value in the range between [a, b]
This means we will take the best predicted value $(.5 * 1 + .5 * .33) =  ~.66.6\%$  and will increase $P(x= increase) * P(\epsilon) = 16. 667\%$ of the time or decrease $P(x= decrease) * P(\epsilon) = 16. 667\%$ of the time. It should be noted that the action value is still clamped between -0.9 and 0.9 regardless of the noise introduced. Alternative exploration policies could be used, but this one was chosen due it’s simplicity, and theoretical capability to explore the whole action space if allowed to run indefinitely.

We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:

\[ [0 0 0 ... 1 ... 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.

The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth. Iroko will generate an adjustment value clamped between $-.9 - .9$. We chose this value range, because we wanted to avoid host starvation.  Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:

\[bandwidth_{t+1}   \leftarrow bandwidth_t +  a_t * bandwidth_t\]

where a is the chosen bandwidth adjustment, t is the previous epoch, and t+1 is the new allocation. This representation could be  modified in a number of ways, such as forcing a minimum expected bandwidth, or add the capacity to starve a host.

>>>>

\subsection{Topology Simulation}
We chose to use  Mininet, a python datacenter emulation API, to prototype Iroko.
Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization
of a datacenter in a single machine ~\cite{mininet}. This allowed us to create a prototype of
Iroko we could test quickly using limited amounts of hardware. Other methods of testing Iroko 
would require more resources, or specialized code specific to the testbed which cannot be 
deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop
Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.

In the Mininet environment, we created a FatTree topology consisting of 20 switches and 16 hosts.
The 20 switches were configured into 5 pods as described in ~\cite{fattree}. This design choice was
to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics
to re-route elephant flows. In ~\cite{hedera}, Hedera was evaluated using a FatTree topology as
we have described, and this allows us to directly compare Iroko against Hedera's performance .

Mininet's lightweight virtualization also allows the API to scale; supporting bigger topologies which otherwise couldn't be prototyped easily ~\cite{mininet}.
This feature will be useful for future research. Using Mininet, we can test Iroko on other topologies beyond the FatTree topology.

@@

In addition, we used the following Linux command:\\
\vspace{1mm}

\texttt{tc -s qdisc show dev <interface>}\\

\vspace{1mm}
where \texttt{<interface>} is the current host we are polling from. Using the 
obtained text, we then applied regex expressions to collect the packet loss, 
packet overlimits, and interface queues. Each of these data points provide further 
information about the current state of the datacenter. Our goal was to 
eliminate dropped packets. We used this technique to help Iroko make rate-limiting 
decisions. Overlimit packets provide information on when a link goes over its 
allocated bandwidth, and adds a potentially useful feature for our controller 
to learn from. Interface queues provide Iroko with information on the queue 
length of the host, which measures how many packets are buffered to be 
processed.  We hoped these additional statistics would provide our controller 
enough data to make meaningful inferences of the appropriate bandwidth 
allocation to each host.

\subsection{Bandwidth Allocation}
Iroko rate limits traffic at the application level, and thus uses UDP instead
of TCP to communicate with hosts. Traffic is generated using a custom load generation
application. We wrote this application to read a traffic matrix shared among all the hosts and send network traffic accordingly.
Iroko manages the current allocated bandwidth per host by sending a controller packet
which updates the current allocated bandwidth.

@@

We chose to use the deep deterministic policy gradient
policy algorithm (DDPG) as described in ~\cite{DDPG} .
DDPG is a form of actor-critic reinforcement learning algorithm ~\cite{Sutton:1998:IRL:551283}
which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor
network gradients relative to the critic's $Q(s,a)$ approximations ~\cite{DDPG}.  Our agent learns using replay memory;
a technique that helps deep reinforcement learning algorithms learn by
making the data uncorrelated via random batch sampling ~\cite{DQLearning}.
This decision might be unnecessary for a datacenter as datacenter traffic may be inherently unpredictable ~\cite{microte}.
Thus we can reasonably assume datacenter traffic is inherently uncorrelated. We recognize these findings
may also give an upper limit to the expected performance of Iroko.

To find the optimal values, an agent must balance exploiting
the current best actions with exploring an environment with 
a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}.
Instead of the exploration approach used in ~\cite{DDPG}, we use $\varepsilon-greedy\ learning$,
where an agent takes the best action with probability $\varepsilon$ and a random action with probability
$1 - \varepsilon$ ~\cite{Sutton:1998:IRL:551283}. Our random action is defined as follows:

@@

uniform(0,1), &\text{P(x = increase) = 1/3},\\
uniform(-1,0), &\text{P(x = decrease) = 1/3},\\
0, & \text{P(x = unchanged) = 1/3}
\end{cases}
\]
\\

where uniform(a,b) refers to taking a random value in the range between [a, b]
This means we will take the best predicted value $(.5 * 1 + .5 * .33) =  ~.66.6\%$  
and will increase $P(x= increase) * P(\varepsilon) = 16. 667\%$ of the time or
decrease $P(x= decrease) * P(\varepsilon) = 16. 667\%$ of the time.
It should be noted that the action value is still clamped between
-0.9 and 0.9 regardless of the noise introduced. Alternative exploration policies 
could be used, but this one was chosen due it's simplicity, and theoretical capability to explore the whole action space if allowed to run indefinitely.

We represent the traffic state of the datacenter as an array combining a bit vector
corresponding to the current host and its allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length. 
The advantage of this technique is that Iroko will learn a policy which considers all hosts together 
in a unified representation.

At each time-step, Iroko increases or decreases the hosts' allocated bandwidth. 
Iroko generates an adjustment value clamped between $-0.9$ and $0.9$. We chose this value range to avoid host starvation. 
Iroko modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:

\[bandwidth_{t+1}   \leftarrow bandwidth_t +  a_t * bandwidth_t\]
\\
where $a$ is the chosen bandwidth adjustment, $t$ is the previous epoch, and $t+1$ is the current epoch. 
This representation could be  modified in a number of ways, for example: enforcing a minimum expected bandwidth, or adding capacity to a starved host.
